---
editor_options: 
  markdown: 
    wrap: 72
---

# Whole-Data Analyses - Classic

This section includes analyses of sustained attention, recognition
memory, and affective rating data from the encoding and retrieval
components of the 'classic' version of the experiment described in
'Experiments'. All participants from the sample were included in the
following analyses. For analyses split by task performance or memory
accuracy, see the 'populations' tab.

```{r , include=FALSE}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
#rm(list = ls())
#load libraries
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(hrbrthemes)
#library(viridis)
library(kableExtra)
library(viridis)
library(lme4)
library(ggpubr)
library(patchwork)
library("gridExtra")
library(sjPlot)

```

```{r , echo=F,include=FALSE}
sust1 <- read_csv("./data_0110/sust1.csv")
sust2 <- read_csv("./data_0110/sust2.csv")
sust3 <- read_csv("./data_0115/sust1.csv")
sust4 <- read_csv("./data_0115/sust2.csv")
sust5 <- read_csv("./data_0116/sust1.csv")
sust6 <- read_csv("./data_0116/sust2.csv")
sust7 <- read_csv("./data_0120/sust1.csv")
sust8 <- read_csv("./data_0120/sust2.csv")
sust9 <- read_csv("./data_0121/sust1.csv")
sust10 <- read_csv("./data_0121/sust2.csv")
sust11 <- read_csv("./data_0130/sust1.csv")
sust12 <- read_csv("./data_0130/sust2.csv")

mem1 <- read_csv("./data_0110/mem1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem2 <- read_csv("./data_0110/mem2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem3 <- read_csv("./data_0115/mem1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem4 <- read_csv("./data_0115/mem2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem5 <- read_csv("./data_0116/mem1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem6 <- read_csv("./data_0116/mem2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem7 <- read_csv("./data_0120/mem1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem8 <- read_csv("./data_0120/mem2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem9 <- read_csv("./data_0121/mem1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem10 <- read_csv("./data_0121/mem2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem11 <- read_csv("./data_0130/mem1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
mem12 <- read_csv("./data_0130/mem2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))

emo1 <- read_csv("./data_0110/emo1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo2 <- read_csv("./data_0110/emo2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo3 <-read_csv("./data_0115/emo1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo4 <-read_csv("./data_0115/emo2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo5 <-read_csv("./data_0116/emo1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo6 <-read_csv("./data_0116/emo2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo7 <-read_csv("./data_0120/emo1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo8 <-read_csv("./data_0120/emo2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo9 <-read_csv("./data_0121/emo1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo10 <-read_csv("./data_0121/emo2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo11 <-read_csv("./data_0130/emo1.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))
emo12 <-read_csv("./data_0130/emo2.csv")%>%
  select(-c("branch-ck1c","randomiser-8rl3","branch-tpdd"))

sust <- rbind( sust1, sust2, sust3, sust4, sust5,sust6,sust7,sust8,sust9,sust10,sust11,sust12)
memory <- rbind(mem1,mem2,mem3,mem4,mem5,mem6,mem7,mem8,mem9,mem10,mem11,mem12)
emo <- rbind(emo1,emo2,emo3,emo4,emo5,emo6,emo7,emo8,emo9,emo10,emo11,emo12)


vamos_aff_s1 <- read_csv("./VAMOS_Set_1_img_info.csv")
vamos_aff_s2 <- read_csv("./NAPS_img_info.csv")



## Filter out the first presentation of the target and only keep responses to the second
 unique_rows <-  sust%>%
  filter(`Spreadsheet: display`=="image")
# Filter rows to keep only those with 'F' if it exists, otherwise keep any row
 filtered_rows <-  unique_rows %>%
  group_by(`Trial Number`,`Participant Private ID`) %>%
  filter(!`Response Type`=="continue")%>%
  filter(if(any(Response == 'F', na.rm = TRUE)) Response == 'F' else TRUE) %>%
  slice(1)%>%
  ungroup()
 filtered_rows <-  filtered_rows %>%
  group_by(`Spreadsheet: img`, `Spreadsheet: type`, `Participant Private ID`) %>%
  mutate(num_rows = n())%>%
  filter(num_rows == 1 | row_number() == 2) %>%
  ungroup() %>%
  select(-num_rows)




 sust_full_all <-  filtered_rows%>%
  arrange(`Participant Private ID`, `Trial Number`)%>%
  mutate(Block_Number = NA,
         Reaction.Time = ifelse(`Response Type` == "timedOut", NA, `Reaction Time`)
  )

 sust_full_all <-  sust_full_all%>%
  group_by(`Participant Private ID`)%>%
  mutate(
    Fitted_Values = ifelse(!is.na(`Reaction Time`), lm(`Reaction Time` ~ `Trial Number`)$fitted.values, NA),
    Detrended_Reaction_Time = `Reaction Time` - Fitted_Values)

 sust_full <- sust_full_all %>%
  filter(!`Spreadsheet: type`=="filler")%>%
  mutate(Block_Number = rep(1:(n() %/% 4 + ifelse(n() %% 4 > 0, 1, 0)), each = 4)[1:n()],
         `Reaction Time` = ifelse(`Response Type` == "timedOut", NA, `Reaction Time`)
  )



 sust_full_window_rt <-  sust_full %>%
  filter(`Spreadsheet: type` == 'window') %>%
  group_by(`Participant Private ID`)%>%
  ungroup()

 sust_full_window_rt <-  sust_full_window_rt %>%
  group_by(`Participant Private ID`,Block_Number)%>%
  mutate(Mean_Reaction_Time = mean(Detrended_Reaction_Time, na.rm = TRUE),
         Mean_Window_1 = mean(Detrended_Reaction_Time, na.rm = TRUE))%>%
  arrange(`Participant Private ID`, `Trial Number`)

 sust_full_target_rt <-  sust_full %>%
  group_by(`Participant Private ID`,Block_Number)%>%
  filter(`Spreadsheet: type` == 'target')



 sust_full_rt <- rbind( sust_full_window_rt, sust_full_target_rt)%>%
  ungroup()%>%
  group_by(`Participant Private ID`)%>%
  mutate(Mean_Window_Scale = scale(Mean_Window_1))%>%
  ungroup()

 sust_full_rt <- sust_full_rt %>%
  arrange(`Participant Private ID`, `Trial Number`)%>%
  group_by(`Participant Private ID`,Block_Number)%>%
  mutate(
    Mean_Reaction_Time = ifelse(`Spreadsheet: type` == "target", mean(Mean_Reaction_Time[`Spreadsheet: type` == "window"]), Mean_Reaction_Time),
    Mean_Window_Scale = ifelse(`Spreadsheet: type` == "target", mean(Mean_Window_Scale[`Spreadsheet: type` == "window"]), Mean_Window_Scale)
  )%>%
  filter(
)


# View the resulting dataframe

#write.csv(sust_full_rt,"sust_full_rt.csv")


 emo <-  emo %>%
  select(`Participant Private ID`, Display, Response, `Spreadsheet: memory`, `Spreadsheet: img`, `Object ID`) %>%
  filter(Display == "image",
         `Spreadsheet: memory` == "target") %>%
  mutate(Response=as.numeric(Response),
         aff_type = ifelse(`Object ID` == "object-51", "valence", 
                           ifelse(`Object ID` == "object-78", "arousal", NA)))%>%
  drop_na()%>%
  filter(Response>=1 & Response<=9)%>%
  rename("Img"="Spreadsheet: img",
         "emo_rate" = "Response")%>%
  group_by(`Participant Private ID`,Img,aff_type) %>%
  filter(row_number() == n() | n() == 1) %>%
  ungroup()%>%
  filter()


#write.csv(emo,"emo.csv")





 memory <-  memory %>%
  filter(Display=="image",
         Response>0,
         Response<7,
         !`Participant Private ID`=="9957478")%>%
  mutate(hits = ifelse(`Spreadsheet: memory` == "target" & (Response == 1 | Response == 2| Response == 3), 1, 0),
         fas = ifelse(`Spreadsheet: memory` == "filler" & (Response == 1 | Response == 2| Response == 3), 1, 0),
         miss = ifelse(`Spreadsheet: memory` == "target" & (Response == 6 |Response == 4 | Response == 5), 1, 0),
         crs = ifelse(`Spreadsheet: memory` == "filler" & (Response == 6 |Response == 4 | Response == 5), 1, 0))
 memory[] <- lapply( memory, function(x) ifelse(x == "A", 1,
                                              ifelse(x == "B", 2,
                                                     ifelse(x == "C", 3,
                                                            ifelse(x == "D", 4,
                                                                   ifelse(x == "E", 5,
                                                                          ifelse(x == "F", 6, x)))))))
 memory <- memory%>%
  group_by(`Participant Private ID`)%>%
  mutate(hit_rate = sum(hits) / (sum(hits)+sum(miss)),
         false_alarm_rate = sum(fas) / (sum(crs)+sum(fas)),
         APrime = ifelse(hit_rate >= false_alarm_rate, 
                         ((hit_rate - false_alarm_rate) * (1 + hit_rate - false_alarm_rate)) / (4 * hit_rate * (1 - false_alarm_rate)) + 0.5, 
                         0.5 - (false_alarm_rate - hit_rate) * (1 + false_alarm_rate - hit_rate) / (4 *false_alarm_rate * (1 - hit_rate))),
         d_prime = (qnorm(hit_rate) - qnorm(false_alarm_rate)) / sqrt(2)
         
  )%>%
  mutate(group=ifelse(`Spreadsheet: memory`=="target","group1","group2"))%>%
  unique()

memory <- memory[memory$APrime>.50,]
sust <- sust[sust$`Participant Private ID` %in% memory[memory$APrime>.50,],]
emo <- emo[emo$`Participant Private ID` %in% memory[memory$APrime>.50,],]
#write.csv(memory, "memory.csv")
 vamos_aff_s1 <- read_csv("./VAMOS_Set_1_img_info.csv")
 vamos_aff_s2 <- read_csv("./NAPS_img_info.csv")


 valence <-  emo[ emo$aff_type=="valence",]
 vamos_aff <- rbind( vamos_aff_s1, vamos_aff_s2)
 
 
 

 mem_only <-  memory%>%
  select(c(`Participant Private ID`,hits,fas,`Spreadsheet: img`,`Trial Number`,Response,`Spreadsheet: memory`,group,APrime))%>%
  filter(`Spreadsheet: memory`=="target")
 mem_only2 <-  memory%>%
  select(c(`Participant Private ID`,hits,fas,`Spreadsheet: img`,`Trial Number`,Response,`Spreadsheet: memory`,group))%>%
  filter(`Spreadsheet: memory`=="filler")

 att_only <-  sust_full_rt%>%
  select(c(`Participant Private ID`,Mean_Reaction_Time,Mean_Window_Scale,`Trial Number`,`Spreadsheet: img`,`Spreadsheet: type`))%>%
  ungroup()%>%
  unique()%>%
  filter(`Spreadsheet: type`== "target")

 mem_att <-merge( mem_only, att_only, by=c("Spreadsheet: img","Participant Private ID"))%>%
  unique()
 mem_att2 <-merge( mem_only2, att_only, by=c("Spreadsheet: img","Participant Private ID"))%>%
  unique()


 mem_att <-  mem_att %>%
  group_by(`Participant Private ID`) %>%
  mutate(zrt = scale(Mean_Reaction_Time))%>%
  rename('Img'= `Spreadsheet: img`)


 mem_only2 <-  mem_only2 %>%
  group_by(`Participant Private ID`) %>%
  rename('Img'= `Spreadsheet: img`)

 emo_mematt <- merge( emo,  mem_att, by= c("Participant Private ID","Img"))
 valence_mematt <-  emo_mematt%>%
  filter(aff_type=="valence")

 arousal_mematt <- emo_mematt%>%
  filter(aff_type=="arousal")

 mem_only_2 <-  mem_only%>%
  rename('Img' =`Spreadsheet: img`)

```

## How often were participants lapsing on the task?

During the go/no-go task (encoding task), participants should have
withheld a response when they saw a second presentation of the target
image. A lapse occurs when they accidentally make a key press on one of
these no-go target images. There were 48 target images shown during this
task for each participant, so the maximum number of lapses is 48. Scroll
through the table to see how often each participant made this mistake.



```{r ,message=F,warning=F}

 rt_analysis <- sust_full_rt%>%
  select(Mean_Reaction_Time,Mean_Window_Scale, Block_Number,`Spreadsheet: type`,Response, `Spreadsheet: img`,`Participant Private ID`)%>%
  group_by(Block_Number,`Participant Private ID`)%>%
  unique()%>%
  mutate(
    targ_perfs = ifelse(`Spreadsheet: type` == "target" & Response == "FALSE", "Lapse", "Correct"),
    Mean_Window_Scale = ifelse(`Spreadsheet: type` == "target", mean(Mean_Window_Scale[`Spreadsheet: type` == "window"]), Mean_Window_Scale)
  )%>%
  filter(`Spreadsheet: type` == "target")
 rt_analysis$NAs = is.na( rt_analysis$targ_perfs)

 rt_analysis <-  rt_analysis%>%
  mutate(
    targ_perf = ifelse(`Spreadsheet: type` == "target" & NAs == "TRUE", "Correct", "Lapse"))%>%
  ungroup()%>%
  group_by(`Participant Private ID`)

 lapse_total <- summarise( rt_analysis, total_lapses = sum(targ_perf == "Lapse"))
    
 lapse_total1 <-  lapse_total %>%
  kable("html") %>%
  kable_styling(full_width = FALSE)%>%
  scroll_box(width="400px",height = "300px") 


 
 d1 <- as.data.frame(lapse_total)
 
  lapse_total1 <-  lapse_total %>%
    filter(total_lapses <22 & total_lapses>0)%>%
  kable("html") %>%
  kable_styling(full_width = FALSE)%>%
  scroll_box(width="400px",height = "300px") 
 lapse_total1
 sust_full_rt_fills <- sust_full_rt %>%
   filter(`Spreadsheet: type`=="filler"|`Spreadsheet: type`=="window")
 
  rt_analysis_fills <- sust_full_rt_fills%>%
  select(Mean_Reaction_Time,Mean_Window_Scale, Block_Number,`Spreadsheet: type`,Response, `Spreadsheet: img`,`Participant Private ID`)%>%
  group_by(Block_Number,`Participant Private ID`)%>%
  unique()%>%
  mutate(
    fills_perfs = ifelse(`Spreadsheet: type` != "target" & Response == "F", "Correct", "Lapse"),
    Mean_Window_Scale = ifelse(`Spreadsheet: type` == "target", mean(Mean_Window_Scale[`Spreadsheet: type` == "window"]), Mean_Window_Scale)
  )
  
 rt_analysis_fills$NAs = is.na( rt_analysis_fills$fills_perfs)

 rt_analysis_fills <-  rt_analysis_fills%>%
  mutate(
    fills_perf = ifelse(`Spreadsheet: type` != "target" & NAs == "TRUE", "Lapse", "Correct"))%>%
  ungroup()%>%
  group_by(`Participant Private ID`)

 lapse_total_fills <- summarise( rt_analysis_fills, total_fills_lapses = sum(fills_perf == "Lapse"))
    
 lapse_total_fills1 <-  lapse_total_fills %>%
   filter(total_fills_lapses<48)%>%
  kable("html") %>%
  kable_styling(full_width = FALSE)%>%
  scroll_box(width="400px",height = "300px") 
 lapse_total_fills1
  
 d2 <- as.data.frame(lapse_total_fills)

too_many_target_lapse <- d1$`Participant Private ID`[d1$total_lapses>24 | d1$total_lapses<0]
too_many_filler_lapse <- d1$`Participant Private ID`[d2$total_fills_lapses>48]
   
   
```



```{r , message=FALSE, warning=FALSE}

 unique_ids <- unique( rt_analysis$`Participant Private ID`)
# Get unique Participant IDs for Correct and Lapse performances
 unique_ids_correct <- unique( rt_analysis$`Participant Private ID`[ rt_analysis$targ_perf == "Correct" & !rt_analysis$`Participant Private ID` %in% too_many_target_lapse & !rt_analysis$`Participant Private ID` %in% too_many_filler_lapse])
 unique_ids_lapse <- unique( rt_analysis$`Participant Private ID`[ rt_analysis$targ_perf == "Lapse"& !rt_analysis$`Participant Private ID` %in% too_many_target_lapse & !rt_analysis$`Participant Private ID` %in% too_many_filler_lapse])

# Get a union of all unique Participant IDs
 all_unique_ids <- union( unique_ids_correct,  unique_ids_lapse)

# Calculate means for Correct performance
 means_by_participant <- tapply(
   rt_analysis$Mean_Window_Scale[ rt_analysis$targ_perf == "Correct"],
   rt_analysis$`Participant Private ID`[ rt_analysis$targ_perf == "Correct"],
  mean,
  na.rm=TRUE
)

# Calculate means for Lapse performance
 means_by_participant2 <- tapply(
   rt_analysis$Mean_Window_Scale[ rt_analysis$targ_perf == "Lapse"],
   rt_analysis$`Participant Private ID`[ rt_analysis$targ_perf == "Lapse"],
  mean,
  na.rm=TRUE
)

# Initialize vectors to store means for all unique Participant IDs
 means_all_correct <- rep(0, length( all_unique_ids))
 means_all_lapse <- rep(0, length( all_unique_ids))
 
matched_indices <- match(names(means_by_participant), all_unique_ids)
valid_indices <- !is.na(matched_indices)

means_all_correct[matched_indices[valid_indices]] <- unname(means_by_participant[valid_indices])

# Replace means for Correct performance with calculated means


 matched_indices2 <- match(names(means_by_participant2), all_unique_ids)
valid_indices2 <- !is.na(matched_indices2)
# Replace means for Lapse performance with calculated means
means_all_lapse[matched_indices2[valid_indices2]] <- unname(means_by_participant2[valid_indices2])

# Convert the result to a data frame
 laps_tot <- data.frame(
  `Participant Private ID` =  all_unique_ids,
  `Correct` =  means_all_correct,
  `Lapse` =  means_all_lapse
)

 lapse_df <- gather( laps_tot, key = "Condition", value = "Mean_Window_Scale", -"Participant.Private.ID")

ggplot( lapse_df, aes(x = Condition, y = Mean_Window_Scale, group = `Participant.Private.ID`)) +
  geom_point(aes(color = factor(`Participant.Private.ID`)), size = 3) +
  geom_line(aes(color = factor(`Participant.Private.ID`)), size = 1) +
  scale_color_viridis(discrete="True", option="A")+
  theme_minimal() +
  xlab("Target Response") +
  ylab(" Mean Window Reaction Time") +
  labs(color = "Participant")+
  theme(legend.position = "none")

library(ggpubr)

t.test(laps_tot$Correct,laps_tot$Lapse,paired = TRUE)

```
<span style="background-color:#A8B89D;">

```{r , warning=F, message=F}

 random_hit_rate <-0.3
 random_fa_rate<-0.3

 chance_d_prime <- (qnorm( random_hit_rate) - qnorm( random_fa_rate))/ sqrt(2)
 
 
 memory <- memory %>%
   filter(!`Participant Private ID` %in% too_many_filler_lapse)%>%
   filter(!`Participant Private ID` %in% too_many_target_lapse)
 
# Summarize data: Sum hits, crs, and count rows (number of trials) for each participant
memory_summary <- memory %>%
  group_by(`Participant Private ID`) %>%
  summarise(
    total_hits = sum(hits),
    total_crs = sum(crs),
    total_success = total_hits + total_crs,  # Sum hits + crs for total success
    n_trials = n()  # Count of rows (number of trials)
  )%>%
  filter(n_trials==96)
p=.5
# Function to apply the binomial test for each participant
apply_binomial_test <- function(total_success, n_trials, p) {
  k <- total_success  # Total successes
  n <- n_trials        # Total trials
  test_result <- binom.test(k, n, p, alternative="greater")  # Perform the binomial test
  return(test_result$p.value)  # Return the p-value
}

# Apply the function to each participant using mapply
memory_summary$p_value <- mapply(apply_binomial_test, 
                                        memory_summary$total_success, 
                                        memory_summary$n_trials,
                                        MoreArgs = list(p = p))

# View the results with the p-values
memory_summary_parts <- memory_summary$`Participant Private ID`[memory_summary$p_value>= 0.05]




memory <- memory%>%
  filter(!`Participant Private ID` %in% memory_summary_parts)



random_hit_rate <-0.3
random_fa_rate<-0.3

chance_d_prime <- (qnorm(random_hit_rate) - qnorm(random_fa_rate))/ sqrt(2)


# Assuming 'memory' is your data frame
memory2 <- memory %>%
  group_by(`Participant Private ID`) %>%
  summarise(total_hits = sum(hits, na.rm = TRUE),
            total_miss = sum(miss, na.rm = TRUE),
            total_fas = sum(fas, na.rm = TRUE), 
            total_crs = sum(crs, na.rm = TRUE))
memory_unique <- memory%>%
  select(APrime,d_prime,`Participant Private ID`)%>%
  unique()



ggplot( memory,aes(x=as.character(`Participant Private ID`),y=APrime,color =as.factor( `Participant Private ID`)))+
  geom_point()+
  geom_hline(yintercept = 0.50, color="red")+
  ylim(0,1)+
  theme_ipsum(axis_title_just = "cc",axis_title_size = 12,axis_text_size = 9)+
  xlab("Participant")+
  ylab("A'")+
  scale_color_viridis(option="A",discrete="True")+
  labs(color = "Participant",
       title="Memory performance (Inclusive)")+
  theme(legend.position="none",axis.text.x = element_blank())

```



```{r ,message=F,warning=F}

   mem_att <- mem_att %>%
  filter(! `Participant Private ID` %in% too_many_filler_lapse) %>%
    filter(! `Participant Private ID` %in% too_many_target_lapse) %>%
  filter(`Participant Private ID` %in% memory$`Participant Private ID`)
mem_att <-merge(mem_att,vamos_aff, by="Img")%>%
  ungroup()%>%
  mutate(zv = scale(Mean_Valence),
         za = scale(Mean_Arousal))%>%
  mutate(z_mem = scale(Memorability_Score))

mem_only2 <-merge(mem_only2,vamos_aff, by="Img")%>%
  ungroup()%>%
  mutate(zv = scale(Mean_Valence),
         za = scale(Mean_Arousal))%>%
  mutate(z_mem = scale(Memorability_Score))

# Fit the GLMM model

mem_att_clean <- mem_att %>%
  filter(!is.na(zrt) & !is.na(hits))

model1 <- glmer(hits ~ zrt + (1 +zrt| `Participant Private ID`), family = binomial, data =  mem_att)
model2 <- glmer(hits ~ z_mem + (1+z_mem | `Participant Private ID`), family = binomial, data =  mem_att)
model3 <-glmer(hits ~ zv + (1 +zv| `Participant Private ID`), family = binomial, data =  mem_att)
model4 <-glmer(hits ~ za + (1 +za| `Participant Private ID`), family = binomial, data =  mem_att)
model5 <-glmer(hits ~ zv + I(zv^2)+(1 +zv| `Participant Private ID`), family = binomial, data =  mem_att)

summary(model1)
summary(model2)
summary(model3)
summary(model4)
summary(model5)
```


```{r}
sustamos_set1_gt <- read.csv("./sustamos_v2_set1.csv")%>%
  rename("Img"="img")
sustamos_set2_gt<- read.csv("./sustamos_v2_set2.csv")%>%
  rename("Img"="img")

set1_full <-merge(sustamos_set1_gt, vamos_aff, by= "Img")%>%
  select(Img,type,Mean_Valence,Mean_Arousal)%>%
  filter(type=="target")%>%
  unique()

set2_full <-merge(sustamos_set2_gt, vamos_aff, by= "Img")%>%
  select(Img,type,Mean_Valence,Mean_Arousal)%>%
  filter(type=="target")%>%
  unique()


emo <- rbind(emo1, emo2) %>%
  select(`Participant Private ID`, Response, `Screen`, `Spreadsheet: img`) %>%
  drop_na() %>%
  mutate(emo_type = ifelse(Screen == "Screen 1", "valence", "arousal")) %>%
  select(-Screen) %>%
  pivot_wider(names_from = "emo_type", values_from = "Response") %>%
  group_by(`Spreadsheet: img`) %>%
  mutate(Mean_Valence2 = mean(as.numeric(valence), na.rm = TRUE),
         Mean_Arousal2 = mean(as.numeric(arousal), na.rm = TRUE))%>%
  rename("Img"=`Spreadsheet: img`)

sets_full <-rbind(set1_full,set2_full)

mem_att2 <-merge(mem_att,emo, by="Img")%>%
  ungroup()%>%
  mutate(zv2 = scale(Mean_Valence2),
         za2 = scale(Mean_Arousal2))

ggplot(mem_att2, aes(x=Mean_Valence,y=Mean_Valence2))+
  geom_point()+
  geom_smooth(method="lm")+
  xlim(1,9)+
  ylim(1,9)

ggplot(mem_att2, aes(x=Mean_Arousal,y=Mean_Arousal2))+
  geom_point()+
  geom_smooth(method="lm")+
  xlim(1,9)+
  ylim(1,9)

mem_att2_sort <- mem_att2%>%
  arrange(desc(Mean_Valence) )%>%
  select(Img,Mean_Valence,Mean_Valence2)%>%
  unique()%>%
  mutate(Img=factor(Img, levels=Img))


mem_att2_long <- mem_att2_sort %>%
  pivot_longer(cols = c(Mean_Valence, Mean_Valence2), 
               names_to = "Valence_Type", 
               values_to = "Valence")

mem_att2_sort2 <- mem_att2%>%
  arrange(desc(Mean_Arousal) )%>%
  select(Img,Mean_Arousal,Mean_Arousal2)%>%
  unique()%>%
  mutate(Img=factor(Img, levels=Img))


mem_att2_long2 <- mem_att2_sort2 %>%
  pivot_longer(cols = c(Mean_Arousal, Mean_Arousal2), 
               names_to = "Arousal_Type", 
               values_to = "Arousal")

# Plot with points and lines
ggplot(mem_att2_long, aes(x = Img, y = Valence, color = Valence_Type, group = Img)) +
  geom_point(size = 3) +  # Adjust size for better visibility
  geom_line() +
  labs(x = "Image", y = "Valence", color = "Valence Type") 

# Plot with points and lines
ggplot(mem_att2_long2, aes(x = Img, y = Arousal, color = Arousal_Type, group = Img)) +
  geom_point(size = 3) +  # Adjust size for better visibility
  geom_line() +
  labs(x = "Image", y = "Arousal", color = "Arousal Type") 

cor.test(mem_att2_long$Valence[mem_att2_long$Valence_Type=="Mean_Valence"],mem_att2_long$Valence[mem_att2_long$Valence_Type=="Mean_Valence2"],method="spearman")

cor.test(mem_att2_long2$Arousal[mem_att2_long2$Arousal_Type=="Mean_Arousal"],mem_att2_long2$Arousal[mem_att2_long2$Arousal_Type=="Mean_Arousal2"],method="spearman")

model6 <-glmer(hits ~ zv2 + (1+zv2 | `Participant Private ID.x`), family = binomial, data =  mem_att2)
model7 <-glmer(hits ~ za2 + (1+za2 | `Participant Private ID.x`), family = binomial, data =  mem_att2)

summary(model6)
summary(model7)
```

```{r,message=FALSE}

memory_rt <- memory %>%
  select(`Reaction Time`,`Participant Private ID`, `Spreadsheet: img`, `Absolute Reaction Time`)%>%
  group_by(`Participant Private ID`)%>%
  mutate(
    max_rt = max(`Reaction Time`)
  )


ggplot(memory_rt, aes(x=as.character(`Participant Private ID`),y=`Reaction Time`,group=as.character(`Participant Private ID`)))+
  geom_violin()+
  geom_hline(yintercept = mean(memory$`Reaction Time`), color="green")+
  ylim(0,30000)



ggplot(memory_rt, aes(x=as.character(`Participant Private ID`),y=`max_rt`,group=as.character(`Participant Private ID`)))+
  geom_point()+
  geom_hline(yintercept = mean(memory$`Reaction Time`), color="green")



# Extract random slopes safely
betas_model1 <- ranef(model1)$`Participant Private ID`[, "zrt", drop = FALSE] %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Participant Private ID")

betas_model2 <- ranef(model2)$`Participant Private ID`[, "z_mem"]
betas_model3 <- ranef(model3)$`Participant Private ID`[, "zv"]
betas_model4 <- ranef(model4)$`Participant Private ID`[, "za"]
betas_model5 <- ranef(model5)$`Participant Private ID`[, "zv"]

# Convert to data frame
betas_df <- data.frame(
  `Participant Private ID` = betas_model1$`Participant Private ID`,
  Beta_Model1 = betas_model1$zrt,
  Beta_Model2 = betas_model2,
  Beta_Model3 = betas_model3,
  Beta_Model4 = betas_model4,
  Beta_Model5 = betas_model5
)

betas_df <-betas_df %>%
  rename("Participant Private ID" = "Participant.Private.ID")
# Assume aprime_df contains A-prime scores
# Example: aprime_df <- data.frame(`Participant Private ID` = unique(mem_att$`Participant Private ID`), aprime = <your_aprime_values>)

betas_df$`Participant Private ID` <-as.numeric(betas_df$`Participant Private ID`)


mem_att_small <- mem_att%>%
  select(`Participant Private ID`,APrime)
# Merge with A-prime scores
merged_df <- betas_df %>%
  merge(mem_att_small, by = "Participant Private ID")

# Plot Beta_Model1 vs A-prime
ggplot(merged_df, aes(x = Beta_Model1, y = APrime)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Beta (Random Slope)", y = "A-prime", title = "Reaction Time") +
  theme_minimal()


# Plot Beta_Model1 vs A-prime
ggplot(merged_df, aes(x = Beta_Model2, y = APrime)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Beta (Random Slope)", y = "A-prime", title = "Memorability") +
  theme_minimal()

# Plot Beta_Model1 vs A-prime
ggplot(merged_df, aes(x = Beta_Model3, y = APrime)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Beta (Random Slope)", y = "A-prime", title = "Valence") +
  theme_minimal()

# Plot Beta_Model1 vs A-prime
ggplot(merged_df, aes(x = Beta_Model4, y = APrime)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Beta (Random Slope)", y = "A-prime", title = "Arousal") +
  theme_minimal()

```
